### 2021-05-24 project_ashish_monograph/CMS.tex
updated

\subsection{CMS detector}
(describe the detector components and what they do) \\

@@

\subsection{The Pixel detector}

@@

\subsection{Cluster reconstruction}
Specific techniques are employed for the reconstruction of clusters in the pixel detector and all the reconstructed clusters are used for counting to calculate the instantaneous luminosity. Algorithm used for finding clusters check for pixels whose signal to noise ratio is more than 6 and then merge adjacent pixel. These algorithms estimates cluster position in two directions and cluster charge. A cluster is a set of adjacent pixels and only pixels above certain minimum value of charge are considered. Position of a cluster in X and Y directions are obtained by fitting X and Y projections to templates that are estimates for cluster shapes \cite{Chatrchyan:2014fea}. \\

>>>>

\subsection{The CMS detector}

@@

\subsection{The CMS Phase I Pixel detector}

@@

\subsection{Cluster reconstruction in pixel detector}
Specific techniques are employed for the reconstruction of clusters in the pixel detector and all the reconstructed clusters are used for counting to calculate the instantaneous luminosity. Algorithm used for finding clusters check for pixels whose signal to noise ratio is more than 6 and then merge adjacent pixel. These algorithms estimates cluster position in two directions and cluster charge. A cluster is a set of adjacent pixels hit by a track originating from interaction point and only pixels above certain minimum value of charge are considered. Position of a cluster in X and Y directions are obtained by fitting X and Y projections to templates that are estimates for cluster shapes \cite{Chatrchyan:2014fea}. \\

### 2021-05-24 project_ashish_monograph/PCC.tex
updated

CMS 2018 data is a collection of runs for each fill in different periods defined during the LHC operation period.

LHC 2018 run is divided into A, B, C and D as described below. \\
\begin{flushleft}
1. Run2018A \\
Run number ranges from 315255 to 316995 \\
Number of runs per period is 146 \\

2. Run2018B\\
Run number ranges from 317080 to 319311\\
Number of runs per period is 132\\

3. Run2018C \\
Run number ranges from 319337 to 320065\\
Number of runs per period is 88 \\

4.Run2018D \\
Run number ranges from 320500 to 323699 \\
Number of runs per period is 262 \\

5. late Run2018D \\
Run number ranges from 320500 to 325175 \\
Number of runs per period is 339 \\
\end{flushleft}

\subsection{Pixel detector Module Selection}
Selection of modules in pixel detector is required because module may experience dynamic inefficiency, bad module chips can stop sending data, chips may not work properly and need reconfiguring. Modules which are observed to function efficiently are considered for luminosity estimation. The module veto lists correspond to late Run2018D that select each detector part barrel layers L2, L3, L4 and forward disks D1, D2, D3 as shown in Fig. 12 and 13. Barrel layer 1 (L1) is not used for luminosity determination \cite{vetolist}. \\

\begin{flushleft}
1. Total number of vetoed modules 1387. \\
2. Number of vetoed modules for L2 1850.\\
3. Number of vetoed modules for L3 1766.\\
4. Number of vetoed modules for L4 1641.\\
5. Number of vetoed modules for D1 1813.\\
6. Number of vetoed modules for D2  1799.\\
7. Number of vetoed modules for D3 1800.\\
\end{flushleft}

>>>>

CMS 2018 data is a collection of runs for each fill in different periods defined during the LHC operation period. LHC 2018 run is divided into A, B, C and D as described below. \\

\begin{flushleft}
1. Run2018A \\
Run number ranges from 315255 to 316995 \\
Number of runs per period is 146 \\
\end{flushleft}

  \begin{flushleft}
2. Run2018B\\
Run number ranges from 317080 to 319311\\
Number of runs per period is 132\\
\end{flushleft}

  \begin{flushleft}
3. Run2018C \\
Run number ranges from 319337 to 320065\\
Number of runs per period is 88 \\
\end{flushleft}

  \begin{flushleft}
4.Run2018D \\
Run number ranges from 320500 to 323699 \\
Number of runs per period is 262 \\
\end{flushleft}

  \begin{flushleft}
5. late Run2018D \\
Run number ranges from 320500 to 325175 \\
Number of runs per period is 339 \\
\end{flushleft}

\subsection{Pixel detector module selection}
Selection of modules in pixel detector is required because module may experience dynamic inefficiency, bad module chips can stop sending data, chips may not work properly and need reconfiguring. Modules which are observed to function efficiently are considered for luminosity estimation. The module veto lists correspond to late Run2018D that select each detector part barrel layers L2, L3, L4 and forward disks D1, D2, D3 as shown in Fig. 12 and 13. Barrel layer 1 (L1) is not used for luminosity determination \cite{vetolist}. \\

\begin{flushleft}
Total number of modules: 1856 \\
BPIX modules per layer:  96 , 224 , 352 , 512\\
FPIX modules per disk:  112 , 112 , 112 , 112 , 112 , 112\\
\end{flushleft}

### 2021-05-24 project_ashish_monograph/bkg.tex
updated

$N_{bg}$ need to be subtracted during vdM scans for precise determination of $\sigma_{vis}$. These contributions to the measured cluster rate are either subtracted before the fit as measured from noncolliding, unpaired bunches or from periods during which the beams were kept at a large distance, or determined in the fit as a constant term added to the fit function as shown in Fig. 13.

@@

Fig. 14 shows the afterglow noise induced by one colliding bunch. Since the distribution is normalized to 1 at the first bin it means the first bin is the colliding bunch. The values at bins greater than 1 are the cluster counts observed for bunches after the colliding bunch divided by the cluster counts observed by the colliding bunch. There are three parts to this distribution: \\

@@

 Afterglow effects for various bunches in a beam is shown in Fig. 15 where the intial rising curve correspond to Type I and decay curve correspond to Type II afterglow effect. It was studied in data taken using random triggers (triggers that fire randomly in bunch crossings spread over the whole LHC orbit except for the abort gap), considering the pixel cluster counts in bunch crossings after bunch trains. The resulting correction depends on the LHC filling scheme and for a typical 2011 fill with 1380 bunches corresponds to a subtraction of close to 2.8$\%$ of the integrated luminosity per lumi section. The afterglow corrections for PCC during 2017 run are in the range 2-5 $\%$ for the Type I corrections and 2-3 $\%$ averaged over all active bunches for the Type II corrections \cite{Sirunyan:2759951}.

\begin{figure}[H]
  \centering
  \includegraphics[width=0.5\columnwidth]{./afterglow.png}
  \caption{Results of the pixel afterglow fit.}
  \label{fig:LHC}
\end{figure}

>>>>

$N_{bg}$ need to be subtracted during vdM scans for precise determination of $\sigma_{vis}$. These contributions to the measured cluster rate are either subtracted before the fit as measured from noncolliding, unpaired bunches or from periods during which the beams were kept at a large distance, or determined in the fit as a constant term added to the fit function as shown in Fig. 15.

@@

Fig. 16 shows the afterglow noise induced by one colliding bunch. Since the distribution is normalized to 1 at the first bin it means the first bin is the colliding bunch. The values at bins greater than 1 are the cluster counts observed for bunches after the colliding bunch divided by the cluster counts observed by the colliding bunch. There are three parts to this distribution: \\

@@

 Afterglow effects for various bunches in a beam is shown in Fig. 17 where the intial rising curve correspond to Type I and decay curve correspond to Type II afterglow effect. It was studied in data taken using random triggers (triggers that fire randomly in bunch crossings spread over the whole LHC orbit except for the abort gap), considering the pixel cluster counts in bunch crossings after bunch trains. The resulting correction depends on the LHC filling scheme and for a typical 2011 fill with 1380 bunches corresponds to a subtraction of close to 2.8$\%$ of the integrated luminosity per lumi section. The afterglow corrections for PCC during 2017 run are in the range 2-5 $\%$ for the Type I corrections and 2-3 $\%$ averaged over all active bunches for the Type II corrections \cite{Sirunyan:2759951}.

\begin{figure}[H]
  \centering
  \includegraphics[width=0.5\columnwidth]{./afterglow.png}
  \caption{Results of the pixel afterglow fit. Type I afterglow effect results from fake electronic pulse response after actual bunch crossing and Type II results from activation of detector material shown as exponentially decaying. }
  \label{fig:LHC}
\end{figure}

### 2021-05-24 project_ashish_monograph/conclusion.tex
progress
updated

\section{Conclusions and outlook} 
\label{sec:conclusion}
sumarize the method and results, and TEPX upgrade work and plans

>>>>

\newpage \section{Summary and Plans} 
\label{sec:conclusion}

In the first part, we processed 2018 CMS data set for period A, B and C to compute luminosity ratios for various subdetectors (L2, L3, L4, D1, D2, D3) of pixel detector. Luminosity determined for pixel subdetectors L3 and F1 shows large instabilities for large lumi section values. We will carry out a detailed study to understand these fluctuations/instabilities observed in the luminosity ratios using new module selections. We will also study vdM background, systematics uncertainties in luminosity measurement and determination of $\sigma_{vis}$ for other luminosity measurement device used during Run 2018 like HFOC.\\

In the second part, we performed a linearity study for Phase II luminosity measurement device Tracker Endcap Pixel Detector (TEPX). Simulated samples from low to high pileup values mimicking Phase II HL-LHC conditions were produced using full CMS detector description and used them to study the behaviour of clusters under different pileup conditions. Different selections on separation between variables defined using cluster position dr and $d\phi$ were applied to study the behaviour of two fold coincidences with pileup to take into account track angle and its curvature. We used different cut values for different TEPX disks and rings to obtain a behaviour closest to linearity for two fold coincidences in $\phi$ and implemented a similar procedure for two fold coincidences in r. New tighter selections dependent on TEPX disks and rings based on truth-match cluster distributions removed physics noise at high pileup. Residual non-linearity for TEPX clusters is with 0.5 \% and for two fold coincidences in $\phi$, r and combined is within 1 \% for all TEPX disks, rings and entire TEPX detector except Ring 5 up to pileup 200.\\

Statistical precision is also calculated for TEPX clusters and coincidences during typical vdM and high pileup physics conditions over different integration time period. For HL-LHC conditions which correspond to pileup 200, the statistical precision which is an important uncertainty for any luminometer is found to be less than 1\%.

### 2021-05-24 project_ashish_monograph/intro.tex
updated

\subsection{Projections for Phase II High Luminosity (HL)-LHC}
LHC during the first run in 2011 and 2012 reached a peak instantaneous luminosity of 0.77 $\times$ $10^{34}$ $cm^{-2}s^{-1}$ which was more than 75$\%$ of its design luminosity and delivered an integrated luminosity of about 30 $fb^{-1}$ to each ATLAS and CMS. The LHC will deliver about 300 $fb^{-1}$ by 2024 \cite{collaborations2019report}. The goal of subsequently run of LHC has been to obtain datasets at higher values of luminosities for precision measurements of the properties of Higgs boson, in order to test the Standard Model pattern of couplings to elementary particles. In order to minimize the machine downtimes and maximize the productive use of the LHC for physics, the replacement of the inner triplet magnets (the one responsible to squeeze the beam at collision) and  of  all  hardware  changes  needed  to  enable  an  ambitious  luminosity  upgrade  will  take  place in parallel during one shutdown, at around 2023-25 (LS3), with some of the modification anticipated in 2019-2020 (LS2). This new phase of the LHC life has been named as High Luminosity LHC (HL-LHC) and has the scope of attaining the astonishing threshold of 3000 $fb^{-1}$ by the year 2036 as shown in Fig. 6. Figure 2 shows a summary of the projected uncertainties in the Higgs boson coupling parameters for Phase II HL-LHC using ggF, VBF, VH, and ${t\bar{t}H}$ production modes where luminosity is one of the dominant uncertainty in Higgs boson coupling measurement \cite{Tanabashi:2018oca}. \\

>>>>

\subsection{Projections for Phase II High Luminosity (HL)-LHC}
LHC during the first run in 2011 and 2012 reached a peak instantaneous luminosity of 0.77 $\times$ $10^{34}$ $cm^{-2}s^{-1}$ which was more than 75$\%$ of its design luminosity and delivered an integrated luminosity of about 30 $fb^{-1}$ to each ATLAS and CMS. The LHC will deliver about 300 $fb^{-1}$ by 2024 \cite{collaborations2019report}. The goal of subsequently run of LHC has been to obtain datasets at higher values of luminosities for precision measurements of the properties of Higgs boson, in order to test the Standard Model pattern of couplings to elementary particles. In order to minimize the machine downtimes and maximize the productive use of the LHC for physics, the replacement of the inner triplet magnets (the one responsible to squeeze the beam at collision) and  of  all  hardware  changes  needed  to  enable  an  ambitious  luminosity  upgrade  will  take  place in parallel during one shutdown, at around 2023-25 (LS3), with some of the modification anticipated in 2019-2020 (LS2). This new phase of the LHC life has been named as High Luminosity LHC (HL-LHC) and has the scope of attaining the astonishing threshold of 3000 $fb^{-1}$ by the year 2036 as shown in Fig. 8. Figure 2 shows a summary of the projected uncertainties in the Higgs boson coupling parameters for Phase II HL-LHC using ggF, VBF, VH, and ${t\bar{t}H}$ production modes where uncertainty in luminosity measurement is one of the dominant uncertainty in Higgs boson coupling measurement \cite{Tanabashi:2018oca}. \\

### 2021-05-24 project_ashish_monograph/results.tex
updated

\begin{figure}[H]
  \centering
  \includegraphics[width=1\columnwidth]{./crop.png}
  \caption{PCC luminosity ($\mu b^{-1}$) and PCC/hfoc luminosity ratio for Run2018A using late Run2018D veto list for pixel detector modules.}
  \label{fig:CMS}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width=1\columnwidth]{./317438(1).png}
  \caption{PCC luminosity ($\mu b^{-1}$) and PCC/hfoc luminosity ratio for Run2018B using late Run2018D veto list for pixel detector modules.}
  \label{fig:CMS}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width=1\columnwidth]{./319659.png}
  \caption{PCC luminosity ($\mu b^{-1}$) and PCC/hfoc luminosity ratio for Run2018C using late Run2018D veto list for pixel detector modules.}
  \label{fig:CMS}
\end{figure}

@@

 \begin{figure}[H]
  \centering
  \includegraphics[width=0.52\columnwidth]{./ls_lumi.png}
  \caption{PCC count as a function of luminosity section (1 lumi section is 23.36s) for Run2018A.}
  \label{fig:CMS}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width=0.5\columnwidth]{./ls_lumi_2018B.png}
  \caption{PCC count as a function of lumi section for Run2018B.}
  \label{fig:CMS}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width=0.5\columnwidth]{./ls_lumi_2018C.png}
  \caption{PCC count as a function of lumi section for Run2018C.}
  \label{fig:CMS}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width=0.52\columnwidth]{./runs.png}
  \caption{PCC integrated luminosity ($\mu b^{-1}$) as a function of run number for Run2018A.}
  \label{fig:CMS}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width=0.52\columnwidth]{./runs_2018B.png}
  \caption{PCC integrated luminosity ($\mu b^{-1}$) as a function of run number for Run2018B.}
  \label{fig:CMS}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width=0.52\columnwidth]{./runs_2018C.png}
  \caption{PCC integrated luminosity ($\mu b^{-1}$) as a function of run number for Run2018C.}
  \label{fig:CMS}
\end{figure}

\subsection{Stability of PCC luminosity for Run 2018 data}

The stability of run 2018 CMS integrated luminosity for Phase I pixel detector calculated using PCC method is investigated by employing late Run2018D veto list and including afterglow effects. New module selections will be determined and used for 2018 luminosity determination. 
\begin{figure}[H]
  \centering
  \includegraphics[width=1\columnwidth]{./t25ld-kgsjf.png}
  \caption{Left: Luminosity ratios for various subdetectors L2, L3, L4, D1, D2, D3 of pixel detector as a function of lumi section for Run2018A. Right: X Profile of luminosity ratios vs lumi section graph for various subdetectors L2, L3, L4, D1, D2, D3 of pixel detector showing luminosity fraction as a function of lumi section (1 lumi section is 23.36s) for Run2018A.}
  \label{fig:CMS}
\end{figure}

>>>>

\begin{figure}[H]
  \centering
  \includegraphics[width=1\columnwidth]{./lhcfill_t.png}
  \caption{Left: Instantaneous luminosity as measured offline by the HFOC luminometer for LHC fill 7270. Right: Ratios of instantaneous luminosity as measured offline by the CMS luminometers as a function of instantaneous luminosity measured by HFOC.}
  \label{fig:CMS}
\end{figure}

@@

 \begin{figure}[H]
  \centering
  \includegraphics[width=1\columnwidth]{./lumiA_merged.png}
  \caption{Left: PCC count as a function of luminosity section (1 lumi section is 23.36s) for Run2018A. Right: PCC integrated luminosity ($\mu b^{-1}$) as a function of run number for Run2018A.}
  \label{fig:CMS}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width=1\columnwidth]{./lumiB_merged.png}
  \caption{Left: PCC count as a function of lumi section for Run2018B. Right: PCC integrated luminosity ($\mu b^{-1}$) as a function of run number for Run2018B}
  \label{fig:CMS}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width=1\columnwidth]{./lumiC_merged.png}
  \caption{Left: PCC count as a function of lumi section for Run2018C. Right: PCC integrated luminosity ($\mu b^{-1}$) as a function of run number for Run2018C.}
  \label{fig:CMS}
\end{figure}

\subsection{Stability of PCC luminosity for Run 2018 CMS data}

The stability of run 2018 CMS integrated luminosity for Phase I pixel detector calculated using PCC method is investigated by employing late Run2018D veto list and including afterglow effects. Luminosity ratios for various subdetectors in pixel detector are computed with respect to total pixel detector luminosity after applying module veto lists for each subdetector. New module selections will be determined and used for 2018 luminosity determination. 
\begin{figure}[H]
  \centering
  \includegraphics[width=1\columnwidth]{./t25ld-kgsjf.png}
  \caption{Left: Luminosity ratios for various subdetectors L2, L3, L4, D1, D2, D3 of pixel detector as a function of lumi section for Run2018A. Right: X Profile of luminosity ratios vs lumi section graph for various subdetectors L2, L3, L4, D1, D2, D3 of pixel detector showing luminosity fraction as a function of lumi section (1 lumi section is 23.36s) for Run2018A.}
  \label{fig:CMS}
\end{figure}

### 2021-05-24 project_ashish_monograph/upgrade.tex
updated

 \section{The TEPX upgrade for HL-LHC}
\label{sec:tepx}

@@

Tracker endcap pixel detector (TEPX) consists of four double disks per side (-Z and +Z) with each double disk containing five rings as shown in Fig. 29 having 20, 28, 36, 44 and 48 modules respectively. One double disk has four surfaces with +Z side containing modules with even module number in front layers (L1 $\&$ L2) and modules with odd module number in back layers (L3 $\&$ L4) from Ring 1 to Ring 4 and for Ring 5, modules with odd module number in front layers and modules with even module number in back layers as shown in Fig. 30. For -Z side, four surfaces contain modules with odd module number in front layer and modules with even module number in back layer from Ring 1 to Ring 4 and for Ring 5, modules with even module number in front layer and modules with odd module number in back layer. \\

\begin{figure}[H]
  \centering
  \includegraphics[width=1 \columnwidth]{./tepx_tt.png}
  \caption{ \onehalfspacing Left: A layout of the CMS Phase-2 inner tracker showing four TEPX disks, eight TFPX disks and four barrel layers. Right: Diagram showing one double disk of TEPX with five rings \cite{Klein:2017nke}.}
  \label{fig:CMS}
\end{figure}

@@

\subsection{Luminosity determination - counting clusters/coincidences}

Instantaneous luminosity determination using PCC method for Phase II HL-LHC will be based on counting the number of clusters or coincidences. The innermost ring of the last disk of TEPX (D4R1) is located at 2.65 m away from the interaction point that is beyond the tracking acceptance ($|\eta| = 4$) and as this region has few tracking points, it can be solely used for the purpose of luminosity measurement by using the full available trigger rate and bandwidth. Two fold coincidences are those hits which are created by the module overlap regions between various layers of one TEPX double disk. Two fold coincidences are better way to distinguish between a real hit and random electrical noise. They are more likely to be real hit than random electrical noise. Two fold coincidence in $\phi$ will involve modules overlapping in the same ring in front and back layers of one double disk as shown in top part of Fig. 31 and Fig. 32 while two fold coincidence in r will require modules overlapping between successive rings in the front (L1 $\&$ L2) and back layers (L3 $\&$ L4) of one double disk as shown in bottom part of Fig. 31. Luminosity calculated based on counting coincidences has an advantage over clusters that afterglow effects are tiny in the case of coincidences  \cite{Collaboration:2706512} \cite{brilsim} \cite{brilsim1}\cite{brilsim2}.\\

\begin{figure}[H]
  \centering
  \includegraphics[width=0.8\columnwidth]{./2foldinrphi.png}
  \caption{ \onehalfspacing Diagram showing modules overlap between the front (L1 \& L2) and back (L3 \& L4) layers of one double disk of TEPX that creates two fold coincidences in $\phi$ and r.}
  \label{fig:CMS}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width=0.6\columnwidth]{./23coin.png}
  \caption{\onehalfspacing Example of two- and threefold coincidence regions on a portion of a single TEPX disk.}
  \label{fig:CMS}
\end{figure}

\subsection{Results: TEPX linearity}

Linearity is one of the systematic uncertainty in the calculation of instantaneous luminosity using PCC and PCC visible cross section $\sigma_{vis}$. A linear relation between the number of clusters and pileup (PU) imply that \\

@@

Linearity indicates that the PCC visible cross section $\sigma_{vis}$ does not depend on the per bunch instantaneous luminosity (ideal luminometer). In ideal scenario, $\sigma_{vis}$ is not dependent on pileup, but this a potential problem with the luminometer. Linearity results for simulated TEPX clusters and coincidences from low to high pileup values are shown from Fig. 33 to Fig. 50. TEPX luminometer proposed for Phase II shows excellent linearity over entire pileup range. A non-linear relation $<N_{cluster}> = \alpha (PU)^{\gamma}, \gamma \neq 1$ would add non-linear terms in the rate equation $R = \sigma_{vis} L_{inst}$  and cause $\sigma_{vis}$ to vary with the per bunch instantaneous luminosity and pileup values.

@@

\begin{figure}[H]
  \centering
  \includegraphics[width=1\columnwidth]{./clustersperringD+4.png}
  \caption{Left: Simulated mean number of clusters for TEPX Disk 4 all rings as a function of pileup. Ring 1 has highest slope and Ring 5 has least slope. Right: Deviation from linearity for clusters for TEPX Disk 4 all rings. Non-linearity is within $1\%$ for all rings over entire pileup range.}
  \label{fig:CMS}
\end{figure}

@@

\subsection{Statistical precision for Phase II luminometer - TEPX}
Statistical precision is another uncertainty that need to be considered for luminosity measurement. It must be kept minimal to achieve 1 $\%$ accuracy for HL-LHC luminosity measurement. \\

@@

Table 1: Expected statistical precision (in $\%$) in head-on collisions during typical vdM conditions with pileup of 0.5 for TEPX two fold coincidences in $\phi$  over different integration time period. 
\begin{center}
    \begin{tabular}{ | l | l | l | p{5cm} |}
    \hline
    vdM (PU 0.5) & Trigger frequency (kHz) & 1 bx, 30 s & 100 bx, 30 s\\ \hline
    TEPX D4R1 2x Coincidences in $\phi$  & 800  & 1.7 & 0.17\\ \hline  
    TEPX 2x Coincidences in $\phi$  & 500  & 0.485 & 0.0485 \\ \hline 
    \hline
    \end{tabular}
\end{center} \\

Table 2: Expected statistical precision (in $\%$) in head-on collisions during physics conditions with pileup of 200 for TEPX two fold coincidences in $\phi$  over different integration time period. 
\begin{center}
    \begin{tabular}{ | l | l | l | p{5cm} |}
    \hline
    Physics (PU 200) & Trigger frequency (kHz) & 1 bx, 4 LN (1.46 s) & 2500 bx, 1 LS (23 s)\\ \hline
    TEPX D4R1 2x Coincidences in $\phi$  & 800  & 0.386 & 0.00193\\ \hline  
    TEPX 2x Coincidences in $\phi$  & 75  & 0.284 & 0.00142\\ \hline 
    \hline
    \end{tabular}
\end{center} \\

Table 3: Expected statistical precision (in $\%$) in head-on collisions during typical vdM conditions with pileup of 0.5 for TEPX clusters and two fold coincidences over different integration time period. 
\begin{center}
    \begin{tabular}{ | l | l | l | p{5cm} |}
    \hline
    vdM (PU 0.5) & Trigger frequency (kHz) & 1 bx, 1 s & 1 bx, 30 s \\ \hline
    TEPX D4R1 clusters  & 1000  & 1.8  & 0.33 \\ \hline  
    TEPX D4R1 2x Coincidences  & 1000  & 5.8 & 1.1 \\ \hline  
    TEPX clusters  & 500  & 0.74  & 0.14\\ \hline
    TEPX 2x Coincidences  & 500  & 2.7 & 0.49\\ \hline
    \hline
    \end{tabular}
\end{center}

Table 4: Expected statistical precision (in $\%$) per second in head-on collisions during physics conditions with pileup of 200 for TEPX clusters and two fold coincidences over different integration time period. 
\begin{center}
    \begin{tabular}{ | l | l | l | p{5cm} |}
    \hline
    Algorithm & Trigger frequency (kHz) & 1 bx, 1 s & 2500 bx, 1 s \\ \hline
    TEPX D4R1 clusters  & 825  & 0.10 & 0.002 \\ \hline  
    TEPX D4R1 2x Coincidences  & 825  & 0.32  & 0.006\\ \hline  
    TEPX clusters  & 75  & 0.095  & 0.002\\ \hline
    TEPX 2x Coincidences  & 75  & 0.35 &0.007 \\ \hline
    \hline
    \end{tabular}
\end{center}

>>>>

 \section{The TEPX upgrade for Phase II HL-LHC}
\label{sec:tepx}

@@

Tracker endcap pixel detector (TEPX) consists of four double disks per side (-Z and +Z) with each double disk containing five rings as shown in Fig. 25 having 20, 28, 36, 44 and 48 modules respectively. One double disk has four surfaces with +Z side containing modules with even module number in front layers (L1 $\&$ L2) and modules with odd module number in back layers (L3 $\&$ L4) from Ring 1 to Ring 4 and for Ring 5, modules with odd module number in front layers and modules with even module number in back layers as shown in Fig. 26. For -Z side, four surfaces contain modules with odd module number in front layer and modules with even module number in back layer from Ring 1 to Ring 4 and for Ring 5, modules with even module number in front layer and modules with odd module number in back layer. \\

\begin{figure}[H]
  \centering
  \includegraphics[width=1 \columnwidth]{./tepx_tt.png}
  \caption{ \onehalfspacing Left: A layout of the CMS Phase II inner tracker showing four TEPX disks, eight TFPX disks and four barrel layers. Right: Diagram showing one double disk of TEPX with five rings \cite{Klein:2017nke}.}
  \label{fig:CMS}
\end{figure}

@@

\subsection{Luminosity determination using TEPX - counting clusters/coincidences}

Instantaneous luminosity determination using PCC method for Phase II HL-LHC will be based on counting the number of clusters in TEPX Disk 4 Ring 1. The innermost ring of the last disk of TEPX (D4R1) is located at 2.65 m away from the interaction point that is beyond the tracking acceptance ($|\eta| = 4$) and as this region has few tracking points, it can be solely used for the purpose of luminosity measurement by using the full available trigger rate and bandwidth.\\

A new algorithm for luminosity determination based on counting two fold coincidences is proposed that applies selections using dr and $d\phi$ variables to minimize non-linearity for high pileup values as these variables take into account track angle and its curvature. Two fold coincidences are those hits which are created by the module overlap regions between various layers of one TEPX double disk. Two fold coincidences are better way to distinguish between a real hit and random electrical noise. They are more likely to be real hit than random electrical noise. Two fold coincidence in $\phi$ will involve modules overlapping in the same ring in front and back layers of one double disk as shown in top part of Fig. 27 and Fig. 28 while two fold coincidence in r will require modules overlapping between successive rings in the front (L1 $\&$ L2) and back layers (L3 $\&$ L4) of one double disk as shown in bottom part of Fig. 27. Luminosity calculated based on counting coincidences has an advantage over clusters that afterglow effects are tiny in the case of coincidences  \cite{Collaboration:2706512} \cite{brilsim} \cite{brilsim1}\cite{brilsim2}.\\

\begin{figure}[H]
  \centering
  \includegraphics[width=0.7\columnwidth]{./2foldinrphi.png}
  \caption{ \onehalfspacing Diagram showing modules overlap between the front (L1 \& L2) and back (L3 \& L4) layers of one double disk of TEPX that creates two fold coincidences in $\phi$ and r.}
  \label{fig:CMS}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width=0.6\columnwidth]{./23coin.png}
  \caption{\onehalfspacing Example of two and threefold coincidence regions on a portion of a single TEPX disk.}
  \label{fig:CMS}
\end{figure}

\subsection{TEPX linearity}

Linearity is one of the systematic uncertainty in the calculation of instantaneous luminosity and PCC visible cross section $\sigma_{vis}$. A linear relation between the number of clusters and pileup (PU) imply that \\

@@

Linearity indicates that the PCC visible cross section $\sigma_{vis}$ does not depend on the per bunch instantaneous luminosity (ideal luminometer). In ideal scenario, $\sigma_{vis}$ is not dependent on pileup, but this a potential problem with the luminometer. Linearity results for simulated TEPX clusters and coincidences from low to high pileup values are shown from Fig. 29 to Fig. 40. TEPX luminometer proposed for Phase II shows excellent linearity over entire pileup range. A non-linear relation $<N_{cluster}> = \alpha (PU)^{\gamma}, \gamma \neq 1$ would add non-linear terms in the rate equation $R = \sigma_{vis} L_{inst}$  and cause $\sigma_{vis}$ to vary with the per bunch instantaneous luminosity and pileup values.

@@

\begin{figure}[H]
  \centering
  \includegraphics[width=1\columnwidth]{./clustersperringD+4.png}
  \caption{Left: Simulated mean number of clusters for +z side TEPX Disk 4 all rings as a function of pileup. Ring 1 has highest slope and Ring 5 has least slope. Right: Deviation from linearity for clusters for TEPX Disk +4 all rings. Non-linearity is within $1\%$ for all rings over entire pileup range.}
  \label{fig:CMS}
\end{figure}

@@

\subsection{Statistical precision for TEPX luminometer}
Statistical precision is another uncertainty that need to be considered for precise luminosity measurement. It must be kept minimal to achieve 1 $\%$ accuracy for HL-LHC luminosity measurement. \\

@@

\newpage
\begin{flushleft} 
  Table 1: Expected statistical precision (in $\%$) in head-on collisions during typical vdM conditions with pileup of 0.5 for TEPX clusters and two fold coincidences over different integration time period.
\end{flushleft}

\begin{center}
\scalebox{1}{
\begin{tabular}{|l | c | c | c |c|}
\hline
vdM (PU 0.5)  & Readout Rate (kHz) &1 bx, 1s & 1 bx, 30s & 100 bx, 30s\\
\hline
TEPXD4R1 Clusters&1000&2.74&0.5 & 0.05\\
\hline
TEPXD4R1 2x Coincidences in phi &1000&8.34&1.52 & 0.152\\
\hline
TEPXD4R1 2x Coincidences in r &1000&16&2.92 & 0.292\\
\hline
TEPXD4R1 2x Coincidences &1000&7.4&1.35 & 0.135\\
\hline
TEPX Clusters&500&0.709&0.129 & 0.0129\\
\hline
TEPX 2x Coincidences in phi &500&2.65&0.485 & 0.0485\\
\hline
TEPX 2x Coincidences in r &500&4.59&0.838 & 0.0838\\
\hline
TEPX 2x Coincidences &500&2.3&0.42 & 0.042\\
\hline
\end{tabular}}
\end{center}

\begin{flushleft} 
Table 2: Expected statistical precision (in $\%$) in head-on collisions during physics conditions with pileup of 200 for TEPX clusters and two fold coincidences over different integration time period.
\end{flushleft}

\begin{center}
\scalebox{1}{
\begin{tabular}{|l | c | c | c |c |}
\hline
 Algorithm & Readout Rate (kHz)& 1 bx, 1s &2500 bx, 1s   & 2500 bx, 1 LS \\
\hline
TEPXD4R1 Clusters&825&0.151&0.00302&0.000624\\
\hline
TEPXD4R1 2x Coincidences in phi &825&0.459&0.00918&0.0019\\
\hline
TEPXD4R1 2x Coincidences in r &825&0.88&0.0176&0.00364\\
\hline
TEPXD4R1 2x Coincidences  &825&0.407&0.00814&0.00169\\
\hline
TEPX Clusters&75&0.0915&0.00183&0.000379\\
\hline
TEPX 2x Coincidences in phi &75&0.343&0.00685&0.00142\\
\hline
TEPX 2x Coincidences in r &75&0.593&0.0119&0.00245\\
\hline
TEPX 2x Coincidences &75&0.297&0.00594&0.00123\\
\hline
\end{tabular}}
\end{center}
